---
title: "Tidymodels Basics"
output: rmarkdown::github_document
---

### Overview
- This notebook covers basic modeling code patterns using the tidymodels framework.
- "The tidymodels framework is a collection of packages for modeling and machine learning using tidyverse principles." - [tidymodels.org](https://www.tidymodels.org/)
- Book: [Tidy Modeling with R by Max Kuhn AND Julia Silge](https://www.tmwr.org/)
- Using OpenIntro Lending Club [loan data](https://www.openintro.org/data/index.php?data=loans_full_schema) uncover key predictors/drivers for loan interest rates.

```{r echo=FALSE, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, dpi=700, fig.cap=TRUE)
```

```{r}
options(scipen=999) 
library(tidyverse)
library(tidymodels)
library(openintro)
library(janitor)
library(skimr)
```

### Data Setup
- Lending Club is a peer to peer loan platform. The data explored in this notebook represents loans made (see link above for more details).
- Imagine we're tasked with understanding what are drivers behind loan interest rates.
- This is a practice scenario (this might not be a relevant business question internal to Lending Club). 

```{r}
individual_loans <- loans_full_schema %>%
      ### build model on subset of individual loans granted
      ### joint also included in the dataset
      filter(application_type=="individual") %>%
      ### subjective human decision to exclude certain features
      dplyr::select(
             -emp_title,
             -emp_length,
             -state,
             -annual_income_joint,
             -verification_income_joint,
             -application_type,
             -grade:-paid_late_fees)

### collect set of columns that contain NAs
### in real world more care even to understand NAs 
### and or use imputation methods to fill NAs in
na_cols <- individual_loans %>%
      summarise_all(.funs = ~sum(is.na(.))) %>%
      gather() %>%
      arrange(desc(value)) %>%
      filter(value>0) %>%
      pull(key)

### drop NA columns to get a practice modeling dataset
individual_loans <- individual_loans %>%
      dplyr::select(-any_of(na_cols))
```

### Brief EDA
- Brief numeric feature summary stats for starting point modeling data.
- More EDA work would be done here in a real world setting to inspire modeling approach and feature engineering. 

```{r}
skimr::skim_without_charts(individual_loans) %>%
      yank("numeric") %>%
      select(-n_missing, -complete_rate)
```
### Create train and test data
```{r}
### generate boiler plate code pattern with usemodels package to speed up workflow
### usemodels::use_glmnet(interest_rate ~ ., data = individual_loans)
### models that usemodels package generates boiler plate code for: ls("package:usemodels", pattern = "^use_")
set.seed(123)
loans_split <- initial_split(individual_loans, prop=0.8, strata=interest_rate)
loans_train <- training(loans_split)
loans_test <- testing(loans_split)

set.seed(123)
### 5 fold cross validation
loans_folds <- vfold_cv(loans_train, v=5)
```

### Set model type
```{r}
lasso_regression_mod <- 
  ### lasso linear regression
  linear_reg(penalty = tune(), mixture = 1) %>% 
  set_mode("regression") %>% 
  set_engine("glmnet") 
```

### Feature engineering recipe
```{r}
lasso_regression_recipe <- 
      recipe(formula = interest_rate ~ ., data = loans_train) %>% 
  step_novel(all_nominal_predictors()) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_nzv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) 
```

### Build modeling workflow
```{r}
lasso_regression_workflow <- 
      workflow() %>% 
      add_model(lasso_regression_mod) %>% 
      add_recipe(lasso_regression_recipe)
```

### Parameter Tuning
- Similar results for small penalty values
```{r}
lasso_regression_grid <- tibble(penalty = 10^seq(-6, -1, length.out = 5)) 

lasso_regression_tune <- 
      tune_grid(lasso_regression_workflow, 
                resamples = loans_folds, 
                grid = lasso_regression_grid)

lasso_regression_tune %>%
      collect_metrics() %>%
      filter(.metric=="rmse")
```
### Fit final model based on param tuning results
```{r}
best_params <- lasso_regression_tune %>% select_best("rmse")

lasso_regression_final_params <- 
  tibble(penalty = best_params$penalty)

final_lasso_regression_workflow <- 
  lasso_regression_workflow %>%
  finalize_workflow(parameters = lasso_regression_final_params)

final_lasso_regression_fit <- 
  final_lasso_regression_workflow %>%
  fit(loans_train)
```

### Variable importance
- Pulse check on features that are most influencial on interest rate prediction

```{r}
final_lasso_regression_fit %>%
  extract_fit_parsnip() %>%
  vip::vi(lambda = best_params$penalty, method="model") %>%
  filter(row_number()<=15) %>%
  ggplot(aes(x = Importance, y = reorder(Variable, Importance), fill = Sign)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)
```

### Test data model performance

```{r}
predict(final_lasso_regression_fit, loans_test) %>%
      mutate(actual = loans_test$interest_rate) %>%
      rmse(truth=actual, estimate=.pred)
```
